---
title: "Data Analysis Exam"
author: "Xinyi Lin"
date: "11/29/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(gee)
library(tidyverse)
library(geepack)
library(nlme)
library(tableone)
```

```{r}
data = read.csv("./ALCDEP.txt", sep = "\t") %>% 
  mutate(Gender = ifelse(Gender==0, "Male", "Female"),
         Gender = as.factor(Gender),
         Treatment = as.factor(Treatment),
         Relapse = as.factor(Relapse))
summary(data)
```

```{r}
# Table1
factorVars <- c("Treatment","Gender","Relapse")
vars <- c("Gender","ND0", "ND30", "ND60", "Relapse")
tableOne <- CreateTableOne(vars = vars, strata = "Treatment", data = data, factorVars = factorVars)
tableOne %>% kableone()
```

## Question 1

In the 30 days following randomization:

$logE[ND30] = \beta_0+ \beta_1Treatment+\beta_2Gender+\beta_3ND0$, where $ND30_i$ follows Poisson distribution.

```{r}
# Table2
glm.model1 = glm(ND30~Treatment+Gender+ND0, family = poisson(), data = data)
summary(glm.model1)
```

```{r}
confint_glm <- function(object, parm, level = 0.95, ...) {
  coef = coef(summary(object)) %>% as.data.frame()
  coef_CI = object %>% confint() %>% as.data.frame()
  table = cbind(coef, coef_CI) %>% 
    mutate(Exp.Est = round(exp(Estimate),4),
           CIL = round(`2.5 %`,4),
           CIU = round(`97.5 %`,4),
           Std.Error = round(`Std. Error`,4),
           Estimate = round(Estimate,4)) %>%
    select(Estimate, Exp.Est, Std.Error, CIL, CIU)
  rownames(table) <- rownames(coef)
  return(table)
}
confint_glm(glm.model1) %>% knitr::kable()
```

According to results above, we can find that the total number of drinks ratio between people who get Treatment 2 and people who get Treatment 1 is `r exp(-0.0156420)` and the total number of drinks ratio between people who get Treatment 3 and people who get Treatment 1 is `r exp(-0.3999763)`. This means for the 30 days following randomization, the effect of Treatment 3 is better than that of Treatment 2 and the effect of Treatment 2 is better than that of Treatment 1.

In the 30 days between the $30^{th}$ and $30^{th}$ day follow-up:

$logE[ND60] = \beta_0+\beta_1Treatment+\beta_2Gender+\beta_3ND0+\beta_4ND30$, where $ND30_i$ follows Poisson distribution.

```{r}
# Table 3
glm.model2 = glm(ND60~Treatment+Gender+ND0 + ND30, family = poisson(), data = data)
summary(glm.model2)
confint_glm(glm.model2) %>% knitr::kable()
```

According to results above, we can find that the total number of drinks ratio between people who get Treatment 2 and people who get Treatment 1 is `r exp(-0.389928)` and the total number of drinks ratio between people who get Treatment 3 and people who get Treatment 1 is `r exp(-0.353110)`. This means for the 30 days following randomization, the effect of Treatment 2 is slightly better than that of Treatment 3 and the effect of Treatment 3 is better than that of Treatment 1.

## Question 2

```{r}
# Figure1
seq_data = data %>% 
  gather(key = "time", value = "ND", ND0:ND60) %>% 
  mutate(time = str_replace(time, "ND", "")) %>% 
  mutate(Gender = as.factor(Gender),
         Treatment = str_c("Treatment",Treatment))
seq_data %>% 
  ggplot(aes(x = time, y = ND, group = sid, color = Treatment)) +
  geom_point()+
  geom_line()+
  facet_grid(. ~ Treatment)
```

Let $y_{ij}$ be the numbers of drinks consumed in the 30 days for subject $i=1,2,...,314$ at visit $j = 0,1,2$. Assume $y_{ij}|b_i$ follows poisson distribution

Model: $log(E[y_{ij}])=\beta_0+\beta_1time_{ij}+\beta_2treatment_{ij} + \beta_3treatment_{ij}*time_{ij} + \beta_4*ND0$.

```{r}
# Table4
spr_data = data %>% 
  gather(key = "time", value = "ND", ND30:ND60) %>% 
  mutate(time = str_replace(time, "ND", "")) %>% 
  mutate(Gender = as.factor(Gender),
         Treatment = as.factor(Treatment))
```

```{r}
gee.model = geeglm(ND ~ Treatment + time + Treatment*time + ND0, id = sid, data = spr_data, family = poisson(), corstr="ar1")
summary(gee.model)
# CI
confint_geeglm <- function(object, parm, level = 0.95, ...) {
    cc <- coef(summary(object))
    mult <- qnorm((1+level)/2)
    citab <- with(as.data.frame(cc),
                  cbind(Estimate = round(Estimate,4),
                        Exp.Est = round(exp(Estimate),4),
                        Std.Error = round(Std.err,4),
                        CIL=round(Estimate-mult*Std.err,4),
                        CIU=round(Estimate+mult*Std.err,4)))
    rownames(citab) <- rownames(cc)
    citab[parm,]
}

confint_geeglm(gee.model) %>% knitr::kable()
```

According to model results, we can find that the ratio of drinks consumed changes between ND30 and ND60 for group in Treatment 2 versus Treatment 1 is significant with value equals to `r exp(-0.373316)` and the ratio of drinks consumed changes between ND30 and ND60 for group in Treatment 3 versus Treatment 1 is not significant with value equals to `r exp(0.004793)`. So the pattern of change in the number of drinks consumed between the Treatment 1 group and Treatment 2 group is difference and the same between the Treatment 1 group and Treatment 3 group.

## Question 3

Model: $logE[ND0] = \beta_0+\beta_1Gender$, where $ND30_i$ follows Poisson distribution.

```{r}
# Table5
glm.model3 = glm(ND0 ~ Gender, data = data, family = poisson())
summary(glm.model3)
confint_glm(glm.model3) %>% knitr::kable()
```

According to above results, we can find that confidence interval of Gender across 0, which means coefficient of Gender is significant with value equals to 0.9975. This means on average, total numder drinks male consumed in baseline 30-day period is `r exp(0.9975)` times compared to that of female.

## Question 4

Let $y_{ij}$ be the numbers of drinks consumed in the 30 days for subject $i=1,2,...,314$ at visit $j = 0,1,2$. Assume $y_{ij}|b_i$ follows poisson distribution

Model: $log(y_{ij})=\beta_0+\beta_1*time_{ij}+\beta_2treatment_{ij} + \beta_3treatment_{ij}*time_{ij} + \beta_4*ND0 + \beta_5*Gender$.

```{r}
# Table6
gee.model2 = geeglm(ND ~ Treatment + time + Treatment*time + ND0 + Gender*Treatment, id = sid, data = spr_data, family = poisson(), corstr="ar1")
summary(gee.model2)
confint_geeglm(gee.model2)%>% knitr::kable()
```

According to results above, we can find that for treatment 1, the log count difference between 

## Question 5

Model $log\frac{P(Relapse=1)}{P(Relapse=0)} = \beta_0+\beta_1Treatment+\beta_2Gender+\beta_3ND0$

```{r}
# Table7
glm.model3 = glm(Relapse ~ Treatment + Gender + ND0, data = data, family = binomial(link = "logit"))
summary(glm.model3)
confint_glm(glm.model3) %>% knitr::kable()
```

As confidence interval of coefficients for treatment 2 and treatment 3 do not across 0, we can conclude that treatment effects difference between treatment 2, treatment 3 versus treatment 1 are significant. The odd ratio of relapse between people in treatment 2 versus people in treatment 1 is `r exp(-1.70055)` and the odd ratio of relapse between people in treatment 2 versus people in treatment 1 is `r exp(-3.36665)`. In general, treatment 3 have best treatment effect on preventing relapse and treatment effect of treatment 2 is better than treatment 1.

## Question 6

```{r}
# delete
seq_data = data %>% 
  gather(key = "time", value = "ND", ND0:ND60) %>% 
  mutate(time = str_replace(time, "ND", "")) %>% 
  mutate(Gender = as.factor(Gender),
         Treatment = str_c("Treatment",Treatment))
seq_data %>% 
  ggplot(aes(x = time, y = ND, group = sid, color = Treatment)) +
  geom_point()+
  geom_line()+
  facet_grid(. ~ Treatment)
```

According to the above plot, we can find that for people in Treatment group 1 and 3, the treatment effects decrease in the 30 days between the $30^{th}$ and $60^{th}$ day follow-up. Total numbers of drinks consumed in the 30 days following randomization are similar to total numbers of drinks consumed in the 30 days between the $30^{th}$ and $60^{th}$ day follow-up. While total numbers of drinks consumed in 30 days keep decrease for people in Treatment group2. According to these trends, treatment 2 might be beneficial once the treatment has stopped.

```{r}
# Figure2
data %>% 
  mutate(Response_p = predict(glm.model3,type="response"),
         Treatment = str_c("Treatment", Treatment)) %>% 
  ggplot(aes(y = Response_p, x = Treatment)) +
  geom_boxplot() +
  scale_y_continuous(name="Relapse Probability")
```

```{r}
# delete
predic_data = data %>% 
  mutate(Response_p = predict(glm.model3,type="response"),
         Treatment = str_c("Treatment", Treatment))
predic_data %>% 
  ggplot(aes(x = Response_p, fill = Treatment)) +
  geom_density(stat = "density") 
```

## Others

Let $\mu_{ij}$ be the average numbers of drinks consumed in the 30 days for treatemnt groups $i=1,2,3$ at visit $j = 0,1,2$.

$H_0:$There is no difference among groups.
$$\mu_{10} = \mu_{20} = \mu_{30}$$
$$\mu_{11} = \mu_{21} = \mu_{31}$$
$$\mu_{12} = \mu_{22} = \mu_{32}$$

$\begin{bmatrix} 1 \ 0 \ -1 \\ 0 \ 1 \ -1 \end{bmatrix} \begin{bmatrix} \mu_{10} \ \mu_{11} \ \mu_{12} \\ \mu_{20} \ \mu_{21} \mu_{22} \\ \mu_{30} \ \mu_{31} \ \mu_{32} \end{bmatrix} I = \begin{bmatrix} 0 \ 0 \ 0 \\ 0 \ 0 \ 0 \end{bmatrix}$

### Question 2

$H_0:$Patterns of change among different groups are the same
$$\mu_{11}-\mu_{10} = \mu_{21}-\mu_{20} = \mu_{31}-\mu_{30}$$
$$\mu_{12}-\mu_{10} = \mu_{22}-\mu_{20} = \mu_{32}-\mu_{30}$$
$$\mu_{12}-\mu_{11} = \mu_{22}-\mu_{21} = \mu_{32}-\mu_{31}$$
$\begin{bmatrix} 1 \ 0 \ -1 \\ 0 \ 1 \ -1 \end{bmatrix} \begin{bmatrix} \mu_{10} \ \mu_{11} \ \mu_{12} \\ \mu_{20} \ \mu_{21} \mu_{22} \\ \mu_{30} \ \mu_{31} \ \mu_{32} \end{bmatrix} \begin{bmatrix} 1 \ 0 \\ -1 \ 1 \\ 0 \ -1 \end{bmatrix} = \begin{bmatrix} 0 \ 0 \ 0 \\ 0 \ 0 \ 0 \end{bmatrix}$

```{r}
library(profileR)
model = pbg(data[,4:6], data$Treatment)
summary(model)
```

## Question 5

Let $y_i$ be whether relapse after 6 months for subject $i=1,2,...,314$. $y_{ij}|b_i$ follows poisson distribution

Model: $log(E[y_{ij}|b_i])=\beta_0+\beta_1*time_{ij}+\beta_2treatment_{ij} + \beta_3treatment_{ij}*time_{ij} + \beta_4*ND0 + \beta_5*Gender$ where $b_i\sim MVN(0,G)$.

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```